Research Proposal

### 2021-10-15 Reports/Journal - Confidence Regions/IT-ConfidenceRegions-V1.tex
Atualizando

\author[1]{Eduarda T.\ C.\ Chagas}
\author[2,3]{Marcelo Queiroz}
\author[4]{Osvaldo A.\ Rosso}
\author[1]{Heitor S.\ Ramos}
\author[2]{Cristopher G.\ S. Freitas}
\author[2]{Leonardo V.\ Pereira}
\author[4]{Alejandro C.\ Frery}

@@

Although very expressive, the Normalized Shannon Entropy is not able to describe all possible underlying dynamics.
In particular, for intermediate values of $H$, there is a wide variety of situations worth characterizing.
To this aim, \citet{LopezRuiz1995} proposed using the disequilibrium  $Q$, a measure of how far $\bm h$ is from equilibrium or non-informative distribution, as the uniform probability distribution function $\bm u$.
%They employed the Euclidean distance between $\bm h$ and the uniform probability distribution function.
It is calculated as:
\begin{equation}
	Q'(\bm{h}, \bm{u}) = \sum_{\ell=1}^{D!} \Big(h(\ell) \log\frac{h(\ell)}{u(\ell)} +
	u(\ell) \log\frac{u(\ell)}{h(\ell)}
	\Big).
\end{equation}
This quantity is also called ``disequilibrium.''
The normalized disequilibrium is $ Q=Q'/\max\{Q'\}$.

>>>>

\author[1]{Eduarda T.\ C.\ Chagas}
\author[2,3]{Marcelo Queiroz}
\author[4]{Osvaldo A.\ Rosso}
\author[1]{Heitor S.\ Ramos}
\author[2]{Cristopher G.\ S. Freitas}
\author[2]{Leonardo V.\ Pereira}
\author[5]{Alejandro C.\ Frery}

@@

Although very expressive, the Normalized Shannon Entropy is not able to describe all possible underlying dynamics.
In particular, for intermediate values of $H$, there is a wide variety of situations worth characterizing.
To this aim, \citet{LopezRuiz1995} proposed using the disequilibrium  $Q$, a measure of how far $\bm h$ is from equilibrium or non-informative distribution, as the uniform probability distribution function $\bm u$.
%They employed the Euclidean distance between $\bm h$ and the uniform probability distribution function.
It is calculated as:
\begin{equation}
	Q'(\bm{h}, \bm{u}) = \sum_{\ell=1}^{D!} \Big(h(\ell) \log\frac{h(\ell)}{u(\ell)} +
	u(\ell) \log\frac{u(\ell)}{h(\ell)}
	\Big).
\end{equation}
This quantity is also called ``disequilibrium.''
The normalized disequilibrium is $ Q=Q'/\max\{Q'\}$, and we chose to work with the uniform probability distribution function as the equilibrium law.
\textcolor{red}{[Eduarda:] alguma referência (dentre as que já citamos) e algumas propriedades?}

### 2021-10-15 Reports/Journal - Confidence Regions/ResearchProposal.tex
update
Atualizando

Our aim is, given a time series $\mathcal X$, computing the point $(h,c)$ and being able to assign a likelihood to the hypothesis that $\mathcal X$ is white noise.
In order to achieve that, we need the distribution of $(\mathfrak{H},\mathfrak{C})$, the bivariate random  variable that produced $(h,c)$, under $H_0:\mathcal X\text { is white noise}$.
We will work with this setup, but later we will relax this null hypothesis.
%
\begin{itemize}
\item[Q1:] What is the distribution of $(\mathfrak{H},\mathfrak{C})$ under $H_0:\mathcal X\text { is white noise}$?
\begin{itemize}
	\item $\bm X = (X_1,X_2,\dots,X_N)$ is white noise
	\item Fix $\tau=1$ and define $D$
	\item $\bm \pi=(\pi_1,\pi_2,\dots,\pi_{n-D+1})$ is the random sequence of patterns from $\bm X$. There are $D!$ possible patterns.
	\item The patterns are not independent because some cannot follow others, but we may shuffle them to avoid such restriction. \textcolor{red}{Eduarda, come up with a proposal for this.}
	\item $\bm h$ is the random histogram of proportions from $\bm \pi$
	\item $(\mathfrak{H},\mathfrak{C})$ is the random point in the Entropy-Complexity plane computed from $\bm h$
	\begin{itemize}	
		\item $H^*$ is the operator that transforms the random vector $\bm h$ into the random variable $\mathfrak{H}$
		\item $C^*$ is the operator that transforms the random vector $\bm h$ in to the random variable $\mathfrak{C}$
		\item Notice that $H^*$ does not compute the entropy of $\bm h$, which is a fixed value; it produces a random variable from the random variable $\bm h$. The same holds for $C^*$; it does not compute the complexity.
	\end{itemize}
\end{itemize}
\item[Q1:] What is the distribution of $
	(\mathfrak{H},\mathfrak{C}) = 
	\big(H^*(\bm h),C^*(\bm h)\big)$? 
\item[$\bullet$] If we stick to $H_0$, $\bm h$ is a collection of random variables taking values in $\{1, 2,\dots, D!\}$ with same probability $1/D!$, but they are not independent.
\item[$\bullet$] We may use the multinomial distribution, but how do we compute the distribution of $H^*(\bm h)$?
\item[$\bullet$] Alternatively, instead of considering such a flexible discrete distribution, we may use a parametric model.
\item[Q2:] We will ask a slightly different question: What is the distribution of $
(\mathfrak{H},\mathfrak{C}) = 
\big(H^*(\bm h),C^*(\bm h)\big)$ when $\bm h\sim\text{dB}(\alpha,\beta,1, D!)$?
\item[] But we go back to a quantity, rather than to a random variable.
\item[Q3:] Assume we have been able to compute the quantities $H^*(\alpha,\beta)$ and $C^*(\alpha,\beta)$ under the $\text{dB}(\alpha,\beta,1, D!)$ model.
\item[] We have the maximum likelihood estimator for $(\alpha,\beta)$, say $(\widehat\alpha,\widehat\beta)$.
\item[] We know that the asymptotic distribution of $(\widehat\alpha,\widehat\beta)$ is a bivariate Gaussian law.
\item[] We may calculate the joint distribution of $\big(H^*(\widehat\alpha, \widehat\beta) , C(\widehat\alpha, \widehat\beta)\big)$.
\end{itemize}
How close are we to finding an approximate distribution for $
(\mathfrak{H},\mathfrak{C})$?
Quite close, I believe.

@@

If $V\sim\text{Beta}(\alpha,\beta)$, then its entropy is
\begin{equation}
H(V) = \log\text{B}(\alpha,\beta) -
(\alpha-1)\psi(\alpha) -(\beta-1) \psi(\beta) +(\alpha+\beta-2) \psi(\alpha+\beta),
\label{Eq:HBeta}
\end{equation}
in which we denoted
$$
\text{B}(\alpha,\beta) = \frac{\Gamma(\alpha)\Gamma(\beta)}{\Gamma(\alpha+\beta)}.
$$
We propose approximating the distribution of $\mathfrak{H}$ by the distribution of 
\begin{equation}
\widehat{H}(V)= \log\text{B}(\widehat\alpha,\widehat\beta) -
(\widehat\alpha-1)\psi(\widehat\alpha) -(\widehat\beta-1) \psi(\widehat\beta) +(\widehat\alpha+\widehat\beta-2) \psi(\widehat\alpha+\widehat\beta),
\end{equation}
which is a transformation of a bivariate Gaussian distribution.

>>>>

Our aim is, given a time series $\mathcal X$, computing the point $(h,c)$ and being able to assign a likelihood to the hypothesis that $\mathcal X$ is white noise.
In order to achieve that, we need the distribution of $(\mathfrak{H},\mathfrak{C})$, the bivariate random  variable that produced $(h,c)$, under $H_0:\mathcal X\text { is white noise}$.
We will work with this setup, but later we will relax this null hypothesis.
%
\begin{itemize}
\item[Q1:] What is the distribution of $(\mathfrak{H},\mathfrak{C})$ under $H_0:\mathcal X\text { is white noise}$?
\begin{itemize}
	\item $\bm X = (X_1,X_2,\dots,X_N)$ is white noise
	\item Fix $\tau=1$ and define $D$
	\item $\bm \pi=(\pi_1,\pi_2,\dots,\pi_{n-D+1})$ is the random sequence of patterns from $\bm X$. There are $D!$ possible patterns.
	\item The patterns are not independent because some cannot follow others, but we may shuffle them to avoid such restriction. \textcolor{red}{Eduarda, come up with a proposal for this.}
	\item $\bm h$ is the random histogram of proportions from $\bm \pi$
	\item $(\mathfrak{H},\mathfrak{C})$ is the random point in the Entropy-Complexity plane computed from $\bm h$
	\begin{itemize}	
		\item $H^*$ is the operator that transforms the random vector $\bm h$ into the random variable $\mathfrak{H}$
		\item $C^*$ is the operator that transforms the random vector $\bm h$ in to the random variable $\mathfrak{C}$
		\item Notice that $H^*$ does not compute the entropy of $\bm h$, which is a fixed value; it produces a random variable from the random variable $\bm h$. The same holds for $C^*$; it does not compute the complexity.
	\end{itemize}
\end{itemize}
\item[Q1:] What is the distribution of $
	(\mathfrak{H},\mathfrak{C}) = 
	\big(H^*(\bm h),C^*(\bm h)\big)$? 
\item[$\bullet$] If we stick to $H_0$, $\bm h$ is a collection of random variables taking values in $\{1, 2,\dots, D!\}$ with same probability $1/D!$, but they are not independent.
\item[$\bullet$] We may use the multinomial distribution, but how do we compute the distribution of $H^*(\bm h)$?
\item[$\bullet$] Alternatively, instead of considering such a flexible discrete distribution, we may use a parametric model.
\item[Q2:] We will ask a slightly different question: What is the distribution of $
(\mathfrak{H},\mathfrak{C}) = 
\big(H^*(\bm h),C^*(\bm h)\big)$ when $\bm h\sim\text{dB}(\alpha,\beta,1, D!)$?
\item[Q3:] Assume we have been able to compute the quantities $H^*(\alpha,\beta)$ and $C^*(\alpha,\beta)$ under the $\text{dB}(\alpha,\beta,1, D!)$ model.
\item[] We have the maximum likelihood estimator for $(\alpha,\beta)$, say $(\widehat\alpha,\widehat\beta)$.
\item[] We know that the asymptotic distribution of $(\widehat\alpha,\widehat\beta)$ is a bivariate Gaussian law.
\item[] We may calculate the joint distribution of $\big(H^*(\widehat\alpha, \widehat\beta) , C(\widehat\alpha, \widehat\beta)\big)$.
\end{itemize}
How close are we to finding an approximate distribution for $
(\mathfrak{H},\mathfrak{C})$?
Quite close, I believe.

@@

If $V\sim\text{Beta}(\alpha,\beta)$, then its entropy is
\begin{equation}
H(V) = \log\text{B}(\alpha,\beta) -
(\alpha-1)\psi(\alpha) -(\beta-1) \psi(\beta) +(\alpha+\beta-2) \psi(\alpha+\beta),
\label{Eq:HBeta}
\end{equation}
in which we denoted
$$
\text{B}(\alpha,\beta) = \frac{\Gamma(\alpha)\Gamma(\beta)}{\Gamma(\alpha+\beta)}.
$$
We propose approximating the distribution of $\mathfrak{H}$ by the distribution of 
\begin{equation}
\widehat{H}(V)= \log\text{B}(\widehat\alpha,\widehat\beta) -
(\widehat\alpha-1)\psi(\widehat\alpha) -(\widehat\beta-1) \psi(\widehat\beta) +(\widehat\alpha+\widehat\beta-2) \psi(\widehat\alpha+\widehat\beta),
\label{eq:Hhat}
\end{equation}
which is a transformation of a bivariate Gaussian distribution.

\textcolor{red}{Magdalena: el estimador de máxima verosimilitud de una transformación del parámetro, es la transformación del estimador de máxima verosimilitud $\widehat{\Psi(\theta)}=\Psi(\widehat\theta)$?}

\textcolor{red}{¿Cuál es la distribución conjuta asintótica de $(\widehat\alpha,\widehat\beta$)? Sabemos que la media es $(\alpha,\beta)$, pero ¿cuál es la matriz de covarianza?}

\textcolor{red}{Cuando la tengamos, ¿podemos resolver~\eqref{eq:Hhat}? Seguro que no. ¿Podemos aproximar la distribución?}
